1.4: Data Visualization with ggplot2

### 2019-09-05 static/slides/04-slides.Rmd
Fix slides 4 final tweak

@@

.center[
![](https://www.dropbox.com/s/f5msffqgvbs4nkb/aes.png?raw=1)
]

@@

# Back to the Original

.left-code[
```{r plot-8a, eval=F, echo=T}
ggplot(data = mpg)+
  aes(x = displ,
        y = hwy)+
  geom_point(aes(color = class))+
  geom_smooth() 
```
]

@@

```{r, fig.retina=3, fig.height=3.5, fig.align="center"}
ggplot(data = mpg)+
  aes(x = displ,
        y = hwy)+
  geom_point(aes(color = class))+
  geom_smooth()+
  facet_wrap(~year) #<<
```
]

@@

```{r, fig.retina=3, fig.height=3.5, fig.align="center"}
ggplot(data = mpg)+
  aes(x = displ,
        y = hwy)+
  geom_point(aes(color = class))+
  geom_smooth()+
  facet_grid(cyl~year) #<<
```
]

@@

```{r, fig.retina=3, fig.align="center"}
ggplot(data = mpg)+
  aes(x = displ,
        y = hwy)+
  geom_point(aes(color = class))+
  geom_smooth()+
  facet_wrap(~year)+
  labs(x = "Engine Displacement (Liters)",
       y = "Highway MPG",
       title = "Car Mileage and Displacement",
       subtitle = "More Displacement Lowers Highway MPG",
       caption = "Source: EPA",
       color = "Vehicle Class")#<<
```
]

@@

class: inverse, center, middle
# Saving Plots

@@

# Saving Plots

# Go Crazy

@@

- **R Studio's [ggplot2 Cheat Sheet](https://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf)**
- **`ggplot2`'s website [reference section](https://ggplot2.tidyverse.org/reference/index.html)**
- STHDA's [be awesome in ggplot2](http://www.sthda.com/english/wiki/be-awesome-in-ggplot2-a-practical-guide-to-be-highly-effective-r-software-and-data-visualization)
- r-statistic's [top 50 ggplot2 visualizations](http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html)
- Hadley Wickham's \href{http://r4ds.had.co.nz/data-visualisation.html}{R for Data Science book chapter on ggplot2}

>>>>

# The Grammar of Graphics (gg): All at Once

All in one command

- Produces plot output in viewer

- Does not save plot
    - Save with `Export` menu in viewer

- Adding layers requires whole code for new plot

.pull-right[
```{r, eval = F}
ggplot(data = mpg)+
  aes(x = displ,
        y = hwy)+
  geom_point()+
  geom_smooth()
```
]

# The Grammar of Graphics (gg): As R Objects

.pull-left[
- Saving as an Object

- Does not show in viewer
    - Rune the name of the plot as command to see it

- Saves your plot as an `R` object

- Can add layers by calling the original plot name
]

```{r, eval=F}
# make and save plot
p <- ggplot(data = mpg)+
  aes(x = displ,
        y = hwy)+
  geom_point()

p # view plot

# add a layer

p + geom_smooth() # shows new plot

p <- p + geom_smooth() # saves and overwrites p
p2 <- p + geom_smooth() # saves as different object

@@

.center[
![](https://www.dropbox.com/s/f5msffqgvbs4nkb/aes.png?raw=1)
]
]
---

@@

# Back to the Original (and saving it)

.left-code[
```{r plot-8a, eval=F, echo=T}
p <- ggplot(data = mpg)+ #<<
  aes(x = displ,
        y = hwy)+
  geom_point(aes(color = class))+
  geom_smooth()

p # show plot
```
]

@@

```{r, fig.retina=3, fig.height=3.5, fig.align="center"}
p + facet_wrap(~year) #<<
```
]

@@

```{r, fig.retina=3, fig.height=3.5, fig.align="center"}
p + facet_grid(cyl~year) #<<
```
]

@@

```{r, fig.retina=3, fig.height=4, fig.align="center"}
p + facet_wrap(~year)+
  labs(x = "Engine Displacement (Liters)", #<<
       y = "Highway MPG", #<<
       title = "Car Mileage and Displacement", #<<
       subtitle = "More Displacement Lowers Highway MPG", #<<
       caption = "Source: EPA", #<<
       color = "Vehicle Class")#<<
```
]

@@

# Go Crazy I

.left-code[
```{r gap-map, echo=F}
library("socviz")
library(mapproj)

us_states <- map_data("state")

election$region <- tolower(election$state)
us_states_elec <- left_join(us_states, election)

theme_map <- function(base_size=9, base_family="") {
    require(grid)
    theme_bw(base_size=base_size, base_family=base_family) %+replace%
        theme(axis.line=element_blank(),
              axis.text=element_blank(),
              axis.ticks=element_blank(),
              axis.title=element_blank(),
              panel.background=element_blank(),
              panel.border=element_blank(),
              panel.grid=element_blank(),
              panel.spacing=unit(0, "lines"),
              plot.background=element_blank(),
              legend.justification = c(0,0),
              legend.position = c(0,0)
              )
}

county_full <- left_join(county_map, county_data, by = "id")
```

```{r gap-map2, eval=F}

# I did some (hidden) data work before this! 
ggplot(data = county_full,
            mapping = aes(x = long, y = lat,
                          fill = pop_dens, 
                          group = group))+ 
  geom_polygon(color = "gray90", size = 0.05)+
  coord_equal()+
  scale_fill_brewer(palette="Blues",
                             labels = c("0-10", "10-50", "50-100", "100-500",
                                        "500-1,000", "1,000-5,000", ">5,000"))+
  labs(fill = "Population per\nsquare mile") +
    theme_map() +
    guides(fill = guide_legend(nrow = 1)) + 
    theme(legend.position = "bottom")
```
]

.right-plot[
```{r ref.label="gap-map2", fig.retina=3, fig.height=5.5, fig.align="center", echo=F}
```
]

@@

# Go Crazy II

@@

- **R Studio's [ggplot2 Cheat Sheet](https://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf)**
- **`ggplot2`'s website [reference section](https://ggplot2.tidyverse.org/reference/index.html)**
- STHDA's [be awesome in ggplot2](http://www.sthda.com/english/wiki/be-awesome-in-ggplot2-a-practical-guide-to-be-highly-effective-r-software-and-data-visualization)
- r-statistic's [top 50 ggplot2 visualizations](http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html)
- Hadley Wickham's[ R for Data Science book chapter on ggplot2](http://r4ds.had.co.nz/data-visualisation.html)
